---
title: "Marginal Model with 12-Month Data"
author: "Jie Zhou"
date: "9/21/2022"
output:  
  workflowr::wflow_html:
    toc: false
---





## clean the data


```{r,message=FALSE,warning=FALSE,results='hide'}
library(gee)
source("./code/datacleaning.r")
index=which(longidata$time==12)
```

## Regression analysis

###  Microbes associated with Bcell
<details>
<summary> Click me </summary>

```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~bcell + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longibcell[index,],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    next()
  }else{
    fm=gee(cbind(y1,y2)~bcell + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longibcell[index,],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=c(r1,r2,r3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}
colnames(rrbcell)=c("coef","pvalue","lower","upper")
fdr=p.adjust(rrbcell[,2],method = "BH")
rrbcell=data.frame(microbe,rrbcell,fdr)
rrbcell=rrbcell[order(rrbcell[,3]),]
rrbcellfdr=rrbcell[which(rrbcell[,6]<0.3),]
```


```{r,echo=FALSE}
rrbcellfdr
```
</details>








###  Microbes associated with CD4
<details>
<summary> Click me </summary>
```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~cd4t + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longicd4t[index,],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    print(k)
    next()
  }else{
    fm=gee(cbind(y1,y2)~cd4t + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longicd4t[index],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=c(r1,r2,r3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}
```
For CD4, GEE method does not yield any significant results.
</details>




###  Microbes associated with CD8
<details>
<summary> Click me </summary>

```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~cd8t + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longicd8t[index,],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    print(k)
    next()
  }else{
    fm=gee(cbind(y1,y2)~cd8t + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longicd8t[index,],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=c(r1,r2,r3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}

```

For CD8, GEE method does not yield any significant results. 
<\details>


###  Microbes associated with Mono
<details> 
<summary> Click me </summary>

```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~mono + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longimono[index],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    next()
  }else{
    fm=gee(cbind(y1,y2)~mono + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longimono[index,],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=c(r1,r2,r3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}

```

```{r,echo=FALSE}
rrbcell
```

</details>

###  Microbes associated with NK
<details>
<summary> Click me </summary>
```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~nk + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longink[index,],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    next()
  }else{
    fm=gee(cbind(y1,y2)~nk + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longink[index,],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=c(r1,r2,r3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}
colnames(rrbcell)=c("coef","pvalue","lower","upper")
fdr=p.adjust(rrbcell[,2],method = "BH")
rrbcell=data.frame(microbe,rrbcell,fdr)
rrbcell=rrbcell[order(rrbcell[,3]),]
rrbcellfdr=rrbcell[which(rrbcell[,6]<0.3),]
```

```{r,echo=FALSE}
rrbcellfdr
```

</details>

###  Microbes associated with NRBC
<details>
<summary> Click me </summary>

```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~nrbc + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longinrbc[index,],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    next()
  }else{
    fm=gee(cbind(y1,y2)~nrbc + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longinrbc[index,],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=c(r1,r2,r3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}
colnames(rrbcell)=c("coef","pvalue","lower","upper")
fdr=round(p.adjust(rrbcell[,2],method = "BH"),3)
rrbcell=data.frame(microbe,rrbcell,fdr)
rrbcell=rrbcell[order(rrbcell[,3]),]
rrbcellfdr=rrbcell[which(rrbcell[,6]<0.3),]
```

```{r,echo=FALSE}
rrbcellfdr
```
</details>


###  Microbes associated with Gran
<details>
<summary> Click me </summary>

```{r,message=FALSE,warning=FALSE,results='hide'}
options(warn = 2)
m=ncol(longidata)-2
n=nrow(longidata)
name=colnames(longidata)[-c(1,2)]
microbe=c()                            
rrbcell=data.frame()
for (k in 1:m) {
  y1=longidata[index,k+2]
  y2=apply(longidata[index,-c(1,2,k+2)],1,sum)
  fm=try({gee(cbind(y1,y2)~gran + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longigran[index,],family="binomial",corstr="independence")},silent = T)
  if (inherits(fm,"try-error")){
    next()
  }else{
    fm=gee(cbind(y1,y2)~gran + bbymale + enrollment_age + imrbthwghtg_all +gestage_all+evercigpreg+alpha+deliver ,id=subject6w, data=longigran[index,],family="binomial",corstr="independence")   
   a=summary(fm)
   r1=a$coefficients[2,1]
   r2=2*(1-pnorm(abs(a$coefficients[2,5])))
   r3= c(r1-1.96*a$coefficients[2,4],r1+ 1.96*a$coefficients[2,4])
   rr=round(c(r1,r2,r3),3)
   microbe=c(microbe,name[k])
    rrbcell=rbind(rrbcell,rr)
     }

}
colnames(rrbcell)=c("coef","pvalue","lower","upper")
fdr=round(p.adjust(rrbcell[,2],method = "BH"),3)
rrbcell=data.frame(microbe,rrbcell,fdr)
rrbcell=rrbcell[order(rrbcell[,3]),]
rrbcellfdr=rrbcell[which(rrbcell[,6]<0.4),]
```
```{r,echo=FALSE}
rrbcellfdr
```

<\details>

